---
title: "Homework 3"
author: "Yueyao Wang"
date: "September 16, 2017"
output: pdf_document
---

### Problem 4  

There are some programming skills in notations, syntax, functions and organisation to make code clean and clear:  

- Using lowercase and underscore in the variable names
- Placing spaces around operators and before left parentheses, except in a function call.
- The opening curly brace should never go on its own line and should always followed by a new line. A closing curly brace should always go on its own line, unless it's followed by else.
- Each lines shouldn't exceed 80 characters.
- When function definition runs over multiple lines, indent the second line where the definition starts.
- Commenting code and use - or = to seperate code into readable chunks.  

I will pay attention to improve the identation and comments.  

### Problem 5  
```{r,eval=FALSE}
library(lintr)
lint(filename = "../02_data_munging_summarizing_R_git/HW2_Wang_Yueyao.Rmd")
```   

- Pay attention that lines shouldn't more than 80 characters.
- Variable and function names should be all lowercase.
- Only use double-quotes for text.
- Commas should always have a space after.

### Problem 6  
```{r}

```


