---
title: 'Dual handling of vector and matrix math'
subtitle: 'Need for speed'
author: "Bob Settlage"
date: '`r Sys.Date()`'
output:
  ioslides_presentation:
    fig_caption: yes
    fig_height: 6
    fig_width: 4
    smaller: yes
  slidy_presentation: default
---

```{r misc_function, eval=T, echo=F, warnings=F}
knitr::opts_chunk$set(echo = F, eval=T, cache=T, tidy.opts=list(width.cutoff=53),
                tidy=T, include=T, message=F, warning=F)
library(ggplot2)
library(ggExtra)
library(MASS)
##this posted answer rocks!!
##https://stackoverflow.com/questions/11022675/rotate-histogram-in-r-or-overlay-a-density-in-a-barplot

scatterBarNorm <- function(x, dcol="blue", lhist=20, num.dnorm=5*lhist, ...){
    ## check input
    stopifnot(ncol(x)==2)
    ## set up layout and graphical parameters
    layMat <- matrix(c(2,0,1,3), ncol=2, byrow=TRUE)
    layout(layMat, widths=c(5/7, 2/7), heights=c(2/7, 5/7))
    ospc <- 0.5 # outer space
    pext <- 4 # par extension down and to the left
    bspc <- 1 # space between scatter plot and bar plots
    par. <- par(mar=c(pext, pext, bspc, bspc),
                oma=rep(ospc, 4)) # plot parameters
    ## scatter plot
    plot(x, xlim=range(x[,1]), ylim=range(x[,2]), pch=20, ...)
    ## 3) determine barplot and height parameter
    ## histogram (for barplot-ting the density)
    xhist <- hist(x[,1], plot=FALSE, breaks=seq(from=min(x[,1]),
                to=max(x[,1]),  length.out=lhist))
    yhist <- hist(x[,2], plot=FALSE, breaks=seq(from=min(x[,2]),
                to=max(x[,2]),  length.out=lhist)) # note: this uses probability=TRUE
    ## determine the plot range and all the things needed for the barplots and lines
    xx <- seq(min(x[,1]), max(x[,1]), length.out=num.dnorm) # evaluation points for the overlaid density
    xy <- dnorm(xx, mean=mean(x[,1]), sd=sd(x[,1])) # density points
    yx <- seq(min(x[,2]), max(x[,2]), length.out=num.dnorm)
    yy <- dnorm(yx, mean=mean(x[,2]), sd=sd(x[,2]))
    ## barplot and line for x (top)
    par(mar=c(0, pext, 0, 0))
    barplot(xhist$density, axes=FALSE, ylim=c(0, max(xhist$density, xy)),
            space=0, col = "grey") # barplot
    lines(seq(from=0, to=lhist-1, length.out=num.dnorm), xy, col=dcol) # line
    ## barplot and line for y (right)
    par(mar=c(pext, 0, 0, 0))
    barplot(yhist$density, axes=FALSE, xlim=c(0, max(yhist$density, yy)),
            space=0, horiz=TRUE, col = "orange") # barplot
    lines(yy, seq(from=0, to=lhist-1, length.out=num.dnorm), col=dcol) # line
    ## restore parameters
    par(par.)
}

```

## Today's Agenda

- Review the apply family of functions  
- Introduce vector and matrix math  
- Talk about timing  
- Discuss methods to make things faster
- Homework 6

## Apply family of functions

We often want to "apply" a function along a "margin" of our data.  In the previous example, we used a function to loop through observers to calculate summary statistics.

In R, we have helper functions to further simplify our code by obviating the for loop.

Apply family:

apply, lapply , sapply, vapply, mapply, rapply, and tapply

Nice tutorial:  
<https://www.r-bloggers.com/r-tutorial-on-the-apply-family-of-functions/>

## Apply detail

*apply(X, MARGIN, FUN, ...)*

```{r echo=T, eval=F, include=T}
    # ?apply
    x <- cbind(x1 = 3, x2 = c(4:1, 2:5))
    dimnames(x)[[1]] <- letters[1:8]
    x
    apply(x, 2, mean, trim = .2)
    col.sums <- apply(x, 2, sum)
    row.sums <- apply(x, 1, sum)
    
    coinToss<-function(x){rbinom(10,1,x)}
    sapply(c(30:40)/100,FUN=coinToss)

```

## What is difference between various apply functions

We could start with the help `?apply, ?sapply`. The main differences are:  
 <http://www.dummies.com/programming/r/how-to-use-the-apply-family-of-functions-in-r/>
 
## Apply functions

```{r apply_descriptions, eval=T, echo=F, include=T, results='asis'}

    library(rvest)
    library(kableExtra)
    webpage <- read_html("http://www.dummies.com/programming/r/how-to-use-the-apply-family-of-functions-in-r/")
    tbls <- html_nodes(webpage, "table")
    tbls_ls <- webpage %>%
            html_nodes("table") %>%
            #.[c(2:5,16:18)] %>%
            html_table(fill = TRUE)
    tbls_ls <- tbls_ls[[1]]
    knitr::kable(tbls_ls) %>% 
        kable_styling(font_size=5)
```

## vectors and matrices in R

Super easy to create:

```{r echo=T, eval=F, include=T}
    a <- c(1:5)
    B <- matrix(1:10,ncol=2)
    a
    B
```

## basic operations

\begin{itemize}
    \item add, multiply vector/matrix by a scalar
    \item add, multiply vectors/matrix
    \item length of vector $$||a|| = \sqrt{\sum_{i=1}^k a_i^2}$$
\end{itemize}

```{r echo=T, eval=F, include=T}

## add scalar
a
a + 5
## multiply by scalar
B
B * 5
## multiply vectors/matrix
a * a
a * B
B * a
## length of vector
sqrt(sum(a*a))
```

## matrix math

<http://www.statmethods.net/advstats/matrix.html>

\begin{itemize}
        \item transpose of vector/matrix
        \item multiply vector/matrix
        \item some special operations
        \begin{itemize}
            \item diag (2 forms)
            \item solve
            \item trace
            \item determinant
            \item Kronecker product
        \end{itemize}
\end{itemize}

```{r echo=T, eval=F, include=T}
## transpose
t(a)
t(t(B))
## matrix multiply
a %*% B
B %*% a
t(B) %*% a
## diag
diag(1,3)
diag(c(1,2,3))
diag(B)
## solve
C <- matrix(1:9 + rnorm(9,0,1e-3),ncol=3) 
C %*% solve(C)
## trace
sum(diag(C))
## det
det(C)
## Kronecker product
I <- diag(3)
J <- matrix(1,3,3)
blockDiag <- kronecker(I, J)
```

## Using duality to do stuff

colMeans.  What if we want the means of a matrix by columns.  Of course there are functions for this, but, for kicks, can we do this with some mixed matrix art ... ??

```{r eval=F, echo=T, include=T}

colMeans(C)
ones <- rep(1,3)
t(ones) %*% C / 3
```

## Making things faster:

A lot of effort has gone into making matrix math faster on computers.  Again, this is a guided tour, so I will show you a few things and hope this sparks your curiousity enough to hit google:

```{r echo=T, eval=F, include=T}

    A = matrix(rnorm(20*40000, mean=0, sd=5), 20, 40000)
    B = matrix(rnorm(20*40000, mean=0, sd=5), 20, 40000)
    system.time({t(a)%*%b})
    system.time({crossprod(a,b)})

```

Don't invert matrices:

<https://www.johndcook.com/blog/2010/01/19/dont-invert-that-matrix/>
<https://www.r-bloggers.com/dont-invert-that-matrix-why-and-how/>

```{r echo=T, eval=F, include=T}

    n <- 5000
    A = matrix(rnorm(n*n, mean=0, sd=5), n, n)
    x <- rnorm(n)
    system.time({b_inverting <- solve(A)%*%x})
    system.time({b_solving <- solve(A,x)})

    max(abs(b_inverting - b_solving))
```


## Homework 6
