---
title: "Homework2"
author: "Yueyao Wang"
date: "September 10, 2017"
output: pdf_document
---

### Problem 4  

Version control can help me keep track of how I coded throughout the whole analysis process. So when I would like to go back to a version of program it would be very easy. This also allows me to do different manipulation on the same dataset at the same time. I don't have to worry about ruining the data. I can also share my work with my teammates when doing the group projects.  

### Problem 5  

####a Sensory Data   

The issue with the data is that the column headers are values, not variable names. "Person1",..."Person5" should be the values of variable person.   

```{r Problem5_Sensory_analysis, message=F,warning=F,echo=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
Sensory_data <- read.table("Sensory.txt",header = F,skip = 1,fill = T,stringsAsFactors = F)
Sensory_tidy1 <- Sensory_data[-1,] %>% filter(V1 %in% 1:10) %>%    
                rename(Item=V1,V1=V2,V2=V3,V3=V4,V4=V5,V5=V6)
Sensory_tidy2 <- Sensory_data[-1,] %>% filter(!(V1 %in% 1:10)) %>% 
                mutate(Item = rep(as.character(1:10),each = 2)) %>% 
                select(Item,V1:V5) %>% 
                mutate(V1 = as.numeric(V1))

Sensory_tidy <- bind_rows(Sensory_tidy1,Sensory_tidy2) %>% 
                gather(person,value,V1:V5) %>% arrange(as.numeric(Item)) %>%
                mutate(person = parse_number(person))

kable(summary(Sensory_tidy), caption="Sensory data summary")
```

####b Long Jump Data  
The issue with this data is that it save the same variable into several columns.  
```{r Problem5_Long_Jump_analysis,message=F,warning=F,echo=FALSE}
LongJump_data <- read.table("LongJumpData.txt",header = T, fill = T) 
LongJump_tidy <- LongJump_data %>% select(Year:Long.2)
##extract two columns
yearcode <- LongJump_tidy[,c(1,3,5,7)] %>% 
            gather(name1, year_code, Year:Year.2)
longjump <- LongJump_tidy[,c(2,4,6,8)] %>% 
            gather(name2,long_jump,Long:Long.2) 
## combine them together and remove missing value
LongJump_tidy <-bind_cols(yearcode,longjump) %>% select(year_code,long_jump)  %>% 
  filter(!is.na(year_code)) %>% mutate(year = year_code + 1990) %>% 
  select(year_code,year,long_jump)

kable(summary(LongJump_tidy), caption="Long Jump data summary")
``` 

####c Brain and Body Weight Data  
The issue with this data is also save value of one variable into serveral columns.   
```{r Problem5_Brain_and_Body_analysis,message=F,warning=F,echo=FALSE}
BrainBody_data <- read.table("BrainandBodyWeight.txt",header = F,fill = T, skip = 1)
colnames(BrainBody_data) <- rep(c("Brain","Body"),times = 3)
BrainBody_tidy <- rbind(BrainBody_data[,1:2],BrainBody_data[,3:4], BrainBody_data[,5:6]) %>% 
                  filter(!is.na(Brain)) %>% arrange(Brain)

kable(summary(BrainBody_tidy), caption="Brain/Body weight data summary")
```


####d Tomato Data  
This data saves multiple data into one cell.  
```{r Problem5_Tomato_analysis,message=F,warning=F,echo=FALSE}
Tomato_data <- read.table("tomato.txt",header = F, fill = T,skip = 2,comment.char = "")
## use number as column name is easy to have error, so combine the number with a character and then can parse the number if needed
colnames(Tomato_data) <- c("variety","C10000","C20000","C30000")
## first gather the header clone into one column
## separate the value column into different replicate and then gather replicates
Tomato_tidy <- Tomato_data %>% gather(clone,value,C10000:C30000) %>% 
                  separate(value,into = paste("replicate",1:3,sep = ""),sep = ",") %>% 
                  gather(replicate,value,replicate1:replicate3) %>%
                  mutate(clone = parse_number(clone), replicate = parse_number(replicate)) %>%
                  select(clone,replicate,value,variety) %>% 
                  arrange(variety,clone,replicate)

kable(summary(Tomato_tidy), caption="Tomato data summary")
```

### Problem 6  
I created a new variable pH_Coverage = pH_Max - pH_Min to combine the information in pH_Min and pH_Max. When I want to measure the relationship between pH and foliage color, I first thought about predicting foliage color using the pH information. However this looks like a multi-dimension classification problem. So I think the other way around and decide to see if we can have a prediction for the pH when we have the foliage color.  
So the model is following:  
Model: $pH\_Coverage = Foliage\_Color + \epsilon$
  
```{r Problem6, message=F,warning=F,echo=FALSE}
library(knitr)
library(swirl)
# Path to data
.datapath <- file.path(path.package('swirl'), 'Courses', 'R_Programming_E', 'Looking_at_Data',
'plant-data.txt')
plants <- read.csv(.datapath, strip.white=TRUE, na.strings="")
# Remove annoying columns
.cols2rm <- c('Accepted.Symbol', 'Synonym.Symbol') 
plants <- plants[, !(names(plants) %in% .cols2rm)]
# Make names pretty
names(plants) <- c('Scientific_Name', 'Duration', 'Active_Growth_Period', 'Foliage_Color', 'pH_Min', 'pH_Max',
                   'Precip_Min', 'Precip_Max',
                   'Shade_Tolerance', 'Temp_Min_F')
# remove observations with missing value
plants <- plants %>% select('Foliage_Color','pH_Min','pH_Max') %>% 
          mutate(pH_Coveragenterval = pH_Max - pH_Min,Foliage_Color = as.factor(Foliage_Color)) %>%
          filter(!is.na(pH_Coveragenterval) & !is.na(Foliage_Color))
lm1 <- lm(formula = pH_Coveragenterval~Foliage_Color+0,data = plants)
kable(summary(lm1)$coefficients)
av1 <- aov(pH_Coveragenterval~Foliage_Color+0,data = plants)
summary(av1)
```  



### Problem 7  
```{r Problem7,message=F,warning=F}
################# load datasets into R   ##################
library(data.table)

#this had the defect code and description
Car_Gebreken_select <- fread(input = "Open_Data_RDW__Gebreken.csv", header = T, select=c(1,6), showProgress=F)
#this has the license plate, inspection date and defect code
Car_Geconstat_select <- fread(input = "Open_Data_RDW__Geconstateerde_Gebreken.csv", header=T, select=c(1,3,5),showProgress=F)
#this has the license plate, make and model of vehicle
Car_Person_select <- fread(input = "Personenauto_basisdata.csv", header=T, showProgress = F, select = c(1,3,4))
Car_Geconstat_select_2017 <- Car_Geconstat_select[grep("2017",Car_Geconstat_select$"Meld datum door keuringsinstantie"),]

################ join them together  ################
### First join Car_Person_select and Car_Geconstat_select by license plate
### Then join the merged one with Car_Gebreken_select by defect code
Merged_Geconstat_Person <- merge(Car_Geconstat_select,Car_Person_select,by = 'Kenteken',all = T)
Merged_All <- merge(Merged_Geconstat_Person,Car_Gebreken_select,by = 'Gebrek identificatie',all = T) 

################# clean the data  ####################
### translate the variables and remove observations containg missing values for any of the variables
colnames(Merged_All) <- c("defect_code","liscence_plate","inspection_date","make","model","defect_description")
### s is a logic vector: whether the observation has any missing value of any of those variables. 
s = apply(Merged_All,MARGIN = 1,function(x) sum(is.na(x)))>0
Merged_tidy <- Merged_All %>% mutate(has_missing_value = s) %>%
              filter(has_missing_value == FALSE)
```  

####d.how many DIFFERENT makes and models of cars you end with (?unique ?distinct ?duplicated) considering only year 2017   
```{r}
Merged_tidy_2017 <- Merged_tidy[grep("2017",Merged_tidy$inspection_date),]
diff_makes <- n_distinct(Merged_tidy_2017$make)
diff_models <- n_distinct(Merged_tidy_2017$model)
```  
The different makes of cars is `r diff_makes` and the different models of cars is `r diff_models`.    

####  e. report a table of the 5 most frequent defects (translated) and the top make/models having that defect (?count) again considering only year 2017   
```{r}
summary1 <- Merged_tidy_2017 %>% group_by(defect_code,make) %>% 
  summarise(make_frequency = n()) %>% group_by(defect_code) %>% 
  mutate(defect_frequency = sum(make_frequency)) %>% 
  mutate(top_make_number = max(make_frequency)) %>%   
  filter(make_frequency %in% top_make_number) %>% 
  arrange(desc(defect_frequency)) %>% 
  select(-top_make_number)
kable(summary1[1:5,],caption = "Top 5 mose frequent defects and the top coresponding make")

```  

####f. use function lm to test for a relationship between number of defects observed by make, report both the coefficient and anova tables (2017 only)   
```{r}
summary2 <- Merged_tidy_2017 %>% group_by(make) %>%
  summarise(number_defect = n_distinct(defect_code))
kable(head(summary2),caption = "number of distinct defects observed by it make")
#fit1 <- lm(number_defect~make,data = summary2)
```

####g.repeat (f) by model (2017 only)  
```{r}
summary3 <- Merged_tidy_2017 %>% group_by(model) %>%
  summarise(number_defect = n_distinct(defect_code))
kable(head(summary3),caption = "number of distinct defects observed by it model")
#fit2 <- lm(number_defect~model,data = summary3)
```

####h.comment on this workflow and how you might be more computationally efficient
I think save those important intermediate dataset for those summarise analysis would improve the efficiency a lot.  

### Appendix: Code  
```{r Appendix, ref.label=c("Problem5_Sensory_analysis","Problem5_Long_Jump_analysis","Problem5_Brain_and_Body_analysis","Problem5_Tomato_analysis","Problem6"), echo=TRUE, eval=F, tidy=TRUE, include=T}
  
```



